Topic 6.2: Open Data

### 2020-04-09 draft_chapter.Rmd
walkthru prep

---
title: "Topic 6.2: Open Data"
author: "Samuel Langton and Reka Solymosi"
date: "1 June 2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

@@

- Topic definition
- Chapter structure

@@

- What is open data?
- What types of organisations release open data?
  - Public sector
  - Private sector
- What are the strengths/limitations of open data?

## Surveying environmental features using Open Street Map

@@

### Downloading Open Street Map Data
- APIs
- Overpass queries and overpass-turbo

@@

osmdata R pacakge.

All queries begin with a bounding box specification to define the study region. This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, but it is generally easier to use a search term. Here, we select Greater London in the United Kingdom using the `getbb()` function, specifying that we want the content as a simple features (sf) object.

@@

We now have our study region defined as the administrative boundaries of Greater London.

@@

Now we have our study region, we can scrape data from the OSM API using the `opq()` function. This allows you to build an Overpass query, outlined in the previous section, from within the R environment. We specify the bounding box defined earlier, and pass this through using a pipe `%>%` to `add_osm_feature()` in which we define what we want to pull from the API. We noted earlier, features recorded on OSM with various tags: keys and values. Here, we specify that we want amenities (the key) defined as bicycle parking (the value). The next function `osmdata_sf()` ensures that the resulting object is a simple features class for easy plotting with `ggplot`. We trim the features pulled from the API based on the Greater London boundaries, otherwise we would end up with information which extends beyond the study region.

```{r}
bikes_sf <- opq(bbox = bb_sf) %>%                                 # select bounding box
  add_osm_feature(key = 'amenity', value = 'bicycle_parking') %>% # select features
  osmdata_sf() %>%                                                # specify class
  trim_osmdata(bb_poly = bb_sf)                                   # trim to bounding box
```

@@

This confirms details like the bounding box, but also provides information on the simple features collected from the query. As one might expect, most information relating to bicyle parking has been recorded using points (i.e. two-dimensional vertices, coordinates) of which we have over seven thousand at the time of writing. We also have around one hundred polygons. For now, let's extract the point information only and then transform the CRS to the BNG.

@@

We can then plot these points over our original boundaries of Greater London, once we have projected it as the BNG too. We reduce the default size of the point to ensure that we avoid too much overlap between bicycle parking locations.

@@

We then need conduct a bit of prliminary data handling: filter crimes which were tagged as bicycle theft, convert the latitude and longitude columns to coordinates with simple features, assign a CRS and then clip the points by our study region. Finally, we transform the CRS to the British National Grid so we can accurately plot these points over our bicycle parking locations.

@@

Then, finally plot everything together, with a few adjustments to the appearance.

@@

- Open topics in open data
- Further applications to crime and place research

@@

- Re-cap
- Wrap-up

>>>>

---
title: "Topic 6.2: Open Data"
author: "Samuel Langton (MMU) & Reka Solymosi (University of Manchester)"
date: "June 2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

@@

- Topic definition.
- Advantages and exciting prospects in open data.
- Chapter structure outline.
- Prerequisites (reference _Geocomputation with R_).

@@

- What is open data?
- What types of organisations release open data?
  - Public sector
  - Private sector
- Why is it needed?
- What are the strengths?
- What are the limitations?
- Substantive examples in criminology (focus on crime and place research).
- Conclude with: a key one is Open Street Map.

## Open Street Map

@@

- Brief history.
- Motivations for and advantages of an open web mapping platform.
- Primary features (keys, values, elements).
- Contributing to the data yourself.

### Downloading Open Street Map Data

- Define and explain APIs.
- For OSM, describe overpass queries and overpass-turbo.
- Query example (brief).

@@

- Made easy through the `osmdata` package in R.
- More straightforward than overpass-turbo.
- Compatible with both `sp` and `sf` classes of spatial data.
- We focus on `sf` as its compatible with the `tidyverse`.
- Outline the key documentation and references.

## Walk-through

First, ensure that you have the relevant packages installed in R. Although the main package used in this demonstration is `osmdata`, for querying the Open Street Map API, we use a number of additional packages for data handling and visualisation. If you don't have these packages installed, use the `install.packages()` function prior to loading each one with `library()`.

```{r, eval = F}
library(osmdata)
library(ggplot2)
library(sf)
library(readr)
library(dplyr)
library(tidyr)
```

All queries begin with a bounding box specification to define the study region. This can be obtained manually, which requires some existing knowledge about an area using the latitude and longitude coordinates, but it is generally easier to use a search term. Here, we select Greater London in the United Kingdom using the `getbb()` function, specifying that we want the content as a simple features (sf) polygon.

@@

We now have our study region defined as the administrative boundaries of Greater London. This can be visualised using `ggplot2` and the _simple features_ geometry `geom_sf()` available with `sf`. Note that by default, the Coordinate Reference System (CRS) is the World Geodetic System 84.

@@

Now we have our study region, we can scrape data from the OSM API using the `opq()` function, which is short for 'Overpass query'. This allows you to build an Overpass query, outlined in the previous section, from within the R environment. We specify the bounding box object which is our study area, and pass this through using a pipe `%>%` to `add_osm_feature()` in which we define what we want to pull from the API. As we noted earlier, features in OSM have are defined through keys and values. Here, we specify that we want amenities (the key) defined as bicycle parking (the value). This query is then piped through to `osmdata_sf()` which ensures that the resulting object is a _simple features_ class for easy plotting with `ggplot2`. We trim the features pulled from the API using `tim_osmdata()` to ensure that everything stays within the boundaries of our study region.

```{r}
bikes_sf <- opq(bbox = bb_sf) %>%                                 # select bounding box
  add_osm_feature(key = 'amenity', value = 'bicycle_parking') %>% # select features
  osmdata_sf() %>%                                                # specify class
  trim_osmdata(bb_poly = bb_sf)                                   # trim to region
```

@@

This confirms details like the bounding box, but also provides information on the simple features collected from the query. As one might expect, most information relating to bicycle parking has been recorded using points (i.e. two-dimensional vertices, coordinates) of which we have over seven thousand at the time of writing. We also have around one hundred polygons. For now, let's extract the point information only and then transform the CRS to the BNG.

@@

We can then plot these points over our original boundaries of Greater London. We reduce the default size of the point to ensure that we avoid too much overlap between bicycle parking locations.

@@

We then need to conduct a bit of preliminary data handling: filter crimes which were tagged as bicycle theft, convert the latitude and longitude columns to coordinates with _simple features_, state the WGS 84 CRS and then clip the points by our study region.

@@

To demonstrate the data in its entirety, we can plot the Greater London boundaries, overlayed with the bicycle parking space locations, and open crime data about bicycle thefts. It is worth clarifying that open police recorded crime data in England and Wales is spatially anonymised by a process of snapping points to a pre-defined grid (see Tompson et al., 2015). For that reason, many of these points overlap, and thus a degree of transparency is used for the points.

@@

- Threats to its sustainability (e.g. licence expiry).
- Prospects in crime of place research.
- Examples of cool projects (e.g. Colouring London).
- Suggestions for new avenues which can expand the field.

@@

- Re-cap on what we've covered.
- Wrap-up the key points.

## References

Tompson, L., Johnson, S., Ashby, M., Perkins, C., & Edwards, P. (2015). UK open source crime data: accuracy and possibilities for research. _Cartography and geographic information science_, 42(2), 97-111.
